Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-08-20 mu/article/article-review7_1.tex
Including all necesary to send the article to PRB

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin
current (PSC) induced by  a linearly polarized light that impinges normally on
the surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional 2D
graphene structures. The first structure, hydrogenated at only one side,
labeled Up, also known as graphone, and the second, labelled Alt, is
25\% hydrogenated at both sides. The hydrogenation opens an energy gap in both
structures. We analyze two possibilities: the first, the spin is fixed
along a chosen direction, and the resulting SVI is calculated.
In the second, we choose the SVI direction along
the surface plane, and calculate the resulting
spin orientation. This is done by changing the energy
$\hbar\omega$ and polarization angle $\alpha$  of the incoming  light. The
results are calculated within a full electronic band structure scheme using the
Density Functional Theory (DFT) in the Local Density Approximation (LDA). The
maxima of the spin-velocities are reached when $\hbar\omega=0.084$\,eV and
$\alpha=35^\circ$ for the Up structure, and $\hbar\omega=0.720$\,eV and
$\alpha=150^\circ$ for the Alt geometry. We find a speed of 668\,Km/s and
645\,Km/s for the Up and the Alt structures, respectively, when the spin points
perpendicularly to the surface. Also, the response is maximized by
fixing the spin-velocity direction 
along a high symmetry  axis, obtaining a speed of 688Km/s
with the spin pointing at $13^\circ$ from the surface normal, for the Up, 
and 906 Km/s and the spin pointing at $60^\circ$ from the surface normal, for
the Alt system. These speed values are  of order of magnitude larger than those of bulk semiconductors,
such as CdSe and GaAs, thus making the hydrogenated graphene structures
excellent candidates for spintronics applications. 
\end{abstract}

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state materials  is
central, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} At present, there is an increasing
interest in attaining the same level of control over the transport of spin at
micro- or nano-scales, as it has been done for the flow of charge in typical
3D-bulk based electronic devices.\cite{awschalomNP2007} Several semiconductor
spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC). One of
the difficulties in creating measurable  spin current and
development of PSC based semiconductor devices is the fact that the spin
relaxation time in conventional semiconducting materials is short to enable the
spin transport, thus resulting in a non-observable spin
current.\cite{murakamiSc03} For PSC there is no net motion of charge; spin-up
electrons move in a given direction, while spin-down electrons travel in the
opposite one. This effect can be due to one-photon absorption of linearly
polarized light by a semiconductor, with filled valence bands and empty
conduction bands, illuminated by light with photon energy larger than the
energy gap. This phenomenon can be due to spin injection,\cite{malPRB03} Hall
Effects,\cite{sinovaPRB04} interference of two optical beams,\cite{bhatPRL00,
najmaiePRB03} or one photon absorption of linearly polarized
light\cite{bhatPRL05}. The last effect has been observed in gallium arsenide
(GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig1}
    \caption{(color online). Top (top panel) and side (bottom panel) views of
the Up structure along with the Cartesian $x$, $y$, and $z$ directions. The
dark (light) spheres are the C (H) atoms. The primitive hexagonal unit cell is
also shown.}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{figures/fig2}
    \caption{(color online). Top (top panel) and side (bottom panel) views of
the Alt structure along with the Cartesian $x$, $y$, and$ z$ directions. The
dark (light) spheres are the C (H) atoms. The primitive rectangular unit cell
is also shown. }
    \label{fig:alt-struc}
\end{figure}

@@

In this section, we  summarize the theoretical approach, involved in
calculation of the spin velocity injection (SVI) resulting from the pure spin
current (PSC).

@@

in this work. To include the coherence, we follow Ref.~\onlinecite{nastosPRB07}

@@

single particle density matrix ${\rho}_{mn}({\mathbf k};t)$, leading to
{\color{red} just for comments to think about, not to introduce the suggestions
right away: above we have $\rho_{mn}$, below $\rho_{cc’}$ time variable has
disappeared below. And $\rho_{cc’}$ contains a and b labels, which are not
Cartesian}

@@

&\frac{\partial \rho_{cc'}({\mathbf k})}{\partial t} =

@@

states, and we take $\epsilon\to 0$ at the end of the calculation. The 
spin-splitting of the valence ($v$) bands is very small, and is neglected throughout
this work.\cite{nastosPRB07} The matrix elements of any operator ${\cal O}$ are

@@

n{\mathbf k}| = 1$, where $n$ goes over all $v$ and $c$ states,
 and the fact that $\rho_{vn}(\mathbf{k})=\rho_{nv}({\mathbf k})=0$ for $n=c,c'$. Therefore, using  Eqs. \eqref{z.4}

@@

–identical to Eq. (3) of Bhat et al.,\cite{bhatPRL05} derived using the

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we calculate the magnitude
of the electron velocity along the plane of the structure, with the spin
polarized along $\hat{\mathbf{b}}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\sigma^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
We also define two special angles
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular to the
incoming light polarization direction, respectively. 
The subscript $\sigma^\mathrm{b}$
denotes the spin along $\hat{\mathbf{b}}$.

@@

Fixing the calculated velocity along $\mathrm{a}=x$ or $\mathrm{a}=y$, we
define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented in the $xyz$ system of
coordinates along the polar angle, defined as
\begin{equation}
\theta_{\mathrm{a}}  (\omega,\alpha) = 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{equation}
and the azimuthal angle
\begin{equation}
\varphi_{\mathrm{a}} (\omega,\alpha) =
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{equation}

@@

We present the calculated results of
$\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ and
$\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ for the Up  and
Alt structures, both noncentrosymmetric 2D carbon systems with
50\% hydrogenation, which are differently structurally arranged. We remind that
the Up structure has hydrogen atoms only on the upper side of the carbon sheet,
while the Alt structure has alternating hydrogen atoms on the upper and bottom
sides. We take the carbon lattice to be along the $xy$ plane for both
structures, and the carbon-hydrogen bonds are outward the plane, they are
perpendicular to  $xz$ plane for the Up structure, as depicted in Figs.
\ref{fig:up-struc}, and off the normal for the Alt structure 
\ref{fig:alt-struc}. The coordinates for the Up and Alt unit cells of the
structures are presented in Tables \ref{tab:up-unitcell} and 
\ref{tab:alt-unitcell}, respectively.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig3}
\caption{(color online).  Spin velocity injection
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ {\it vs.} photon energy
  $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The largest
  velocity are at the low energy regions of the spectra for the Alt and Up
  structures, becoming different from zero at the energy gap of each structure.
  In the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to
  the 3D case of CdSe and GaAs, shown at the bottom panel.}
\label{fig:vab-str-comp}
\end{figure}

In Fig. \ref{fig:vab-str-comp}, we show $\mathcal{V}^{\mathrm{ab}}
(\omega,\alpha)$ {\it vs.} $\hbar\omega$ for the 
velocity and spin
directions
$\hat{\mathbf{a}}$ and $\hat{\mathbf{b}}$, and for the angle $\alpha$, at which
the signal is maximized, for the Up  and Alt structures, and for CdSe and GaAs
bulk systems, shown for comparison. As expected from
Eq.~\eqref{eq:mu}, 
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ starts rising 
from zero right at the
corresponding energy gap of each system. For the 2D structures considered, the
spectrum contains two narrow energy regions with strong response, while for the
bulk systems, the spectra covers a rather wide energy range, but with a much
weaker response. For the Up structure, at $\mathrm{a}\mathrm{b}=yz$ and
$\alpha=35^\circ$ the response is maximized, which means that an incoming light
with its electric field polarized at 35$^\circ$ from the $x$ direction will
induce electrons to move along $y$, parallel to the surface, with their spin
polarized along $z$, perpendicular to the surface. Right at the energy onset, ${\cal
V}^{yz}(\omega,\alpha)=668$\,Km/s, remains almost constant for 65 meV, and then
decreases to zero. A second region with high velocity is above 1.946 eV with
two, opposite in sign, maxima of the speed: ${\cal
V}^{yz}(\omega,\alpha)=266.3$ Km/s at $\hbar\omega=1.954$\,eV, and ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)=-241.4$ Km/s at
$\hbar\omega=1.958$\,eV; a positive (negative) ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ means that the electrons move
parallel (antiparallel) to the electric field. Likewise, for the Alt structure,
we also find that $\mathrm{a}\mathrm{b}=yz$ and $\alpha=150^\circ$ maximizes
the response, where two extreme values of ${\cal V}^{\mathrm{y}\mathrm{z}}$ are
found, one at  $\hbar\omega=0.720$\,eV of ${\cal
V}^{\mathrm{y}\mathrm{z}}=-711.9$ Km/s, and the other at
$\hbar\omega=0.911$\,eV of ${\cal V}^{\mathrm{y}\mathrm{z}}=-330.6$ Km/s.

For the bulk structures, we calculate ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$
from Eq.~\eqref{eq:vab-w} by simply using $\boldsymbol{\mu}_{\mathrm{max}}$.
For CdSe, we find that for $\hbar\omega=0.844$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to \mu^{zzzz}$, and ${\cal
V}^{\mathrm{z}\mathrm{z}}(\omega)=-59.0$ Km/s, and for GaAs at
$\hbar\omega=2.324$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to\mu^{\mathrm{aaaa}}$ and ${\cal
V}^{\mathrm{a}\mathrm{a}}(\omega,\alpha)=-28.7$\,Km/s, with $\mathrm{a}=x,y,z$.
For these bulk semiconductors, the $x$, $y$, and $z$ axis are taken along the
standard cubic unit cell directions, [100], [010], and [001], respectively. In
Table \ref{tab:vab-str-comp}, we compare ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ for the 2D structures considered and
bulk crystals. We stress that, as shown in the figure, for the 2D structures
have maxima in ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$ are higher than
for the bulk crystals more than order of magnitude. In particular, the Alt structure demonstrates a ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$ about 12 times larger than that
of CdSe and GaAs.

{\color{red} Finished here; A.}
\begin{table}%[b]
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
System \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$}\\
\cline{5-6}
& type & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
Up  & 2D   & 35    & 0.084  & $\mathrm{yz}$ &  660.5 \\
      &      &       & 1.954  & $\mathrm{yz}$ &  266.3 \\
      &      &       & 1.958  & $\mathrm{yz}$ & -241.4 \\
Alt & 2D   & 150   & 0.720  & $\mathrm{yz}$ & -711.9 \\
      &      &       & 0.911  & $\mathrm{yz}$ & -330.6 \\
CdSe  & bulk & --    & 0.844  & $\mathrm{zz}$ &  -59.0 \\
GaAs  & bulk & --    & 2.324  & $\mathrm{xx}$ &  -28.7 \\
\hline
\end{tabular}
\caption{Comparison of the reported maximum values of
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ for the different structures and
their corresponding polarization angle $\alpha$ and energy $\hbar\omega$ . }
\label{tab:vab-str-comp}
\end{table}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig4}
\caption{(color online). Color online. For the Up structure, the top panel
shows ${\cal V}_{\sigma^\mathrm{z}}(\omega,\alpha)$ {\it vs.} $\hbar\omega$ and
$\alpha$, and the bottom panel shows
$\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ (right scale, red line), and ${\cal
V}_{\sigma^\mathrm{z}}(\omega,\alpha)$ (left scale, black solid line), {\it
vs.} $\alpha$, for $\hbar\omega=0.084$\,eV, i.e. along the ridge shown in the
3D plot. }
\label{fig:up-vsz-w1}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig5}
\caption{(color online). For the Up structure, the top panel shows ${\cal
V}_{\sigma^\mathrm{z}} (\omega,\alpha)$ {\it vs.} $\hbar\omega$ and $\alpha$,
and the bottom panel shows $\gamma_{\sigma^\mathrm{z}} (\omega,\alpha)$ (right
scale, red line, long dashes), and ${\cal V}_{\sigma^\mathrm{z}}
(\omega,\alpha)$ (left scale, black solid line), {\it vs.} $\alpha$, for
$\hbar\omega=1.954$\,eV, i.e. along the highest ridge shown in the 3D plot. }
\label{fig:up-vsz-w2}
\end{figure}

@@

Here we calculated $\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ (Eq. 
\eqref{eq:vv-mag}) after fixing the electron velocity direction, $\hat{\mathbf{a}}$,
to the $x$ or $y$ direction along the surface of the Up and Alt structures.
From Eqns. \eqref{eq:polar-ang} and \eqref{eq:azimuthal-ang}, 
we
determined the corresponding polar angle, $\theta_{\mathrm{a}}
(\omega,\alpha)$, and azimuthal angle, $\varphi_{\mathrm{a}} (\omega,\alpha)$.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{figures/fig7}
\caption{(color online). For the Up structure we show the velocity
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$ (left scale, black line), the polar
angle $\theta_{\mathrm{a}} (\omega,\alpha)$ (right scale, red line), and  the
azimuthal angle $\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue
line), {\it vs.} $\hbar\omega$, for $\alpha=35^\circ$, and $\mathrm{a}=x$ and
$\mathrm{a}=y$.}
\label{fig:up-vab-comp-rtp-1}
\end{figure}

>>>>

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin
current (PSC) induced by  a linearly polarized light that impinges normally on
the surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional 2D
graphene structures. The first structure, hydrogenated at only one side,
labeled Up, also known as graphone, and the second, labelled Alt, is
25\% hydrogenated at both sides. The hydrogenation opens an energy gap in both
structures. We analyze two possibilities: in the first, the spin is fixed
along a chosen direction, and the resulting SVI is calculated.
In the second, we choose the SVI direction along
the surface plane, and calculate the resulting
spin orientation. This is done by changing the energy
$\hbar\omega$ and polarization angle $\alpha$  of the incoming  light. The
results are calculated within a full electronic band structure scheme using the
Density Functional Theory (DFT) in the Local Density Approximation (LDA). The
maxima of the spin-velocities are reached when $\hbar\omega=0.084$\,eV and
$\alpha=35^\circ$ for the Up structure, and $\hbar\omega=0.720$\,eV and
$\alpha=150^\circ$ for the Alt geometry. We find a speed of 668\,Km/s and
645\,Km/s for the Up and the Alt structures, respectively, when the spin points
perpendicularly to the surface. Also, the response is maximized by
fixing the spin-velocity direction 
along a high symmetry  axis, obtaining a speed of 688Km/s
with the spin pointing at $13^\circ$ from the surface normal, for the Up, 
and 906 Km/s and the spin pointing at $60^\circ$ from the surface normal, for
the Alt system. These speed values are  of order of magnitude larger than those of bulk semiconductors,
such as CdSe and GaAs, thus making the hydrogenated graphene structures
excellent candidates for spintronics applications. 
\end{abstract}

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of the electron spin in a solid state materials  is
central, adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} At present, there is an increasing
interest in attaining the same level of control over the transport of spin at
micro- or nano-scales, as it has been done for the flow of charge in typical
3D-bulk based electronic devices.\cite{awschalomNP2007} Several semiconductor
spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC). One of
the difficulties in creating measurable  spin current and
development of PSC based semiconductor devices is the fact that the spin
relaxation time in conventional semiconducting materials is short to enable the
spin transport, thus resulting in a non-observable spin
current.\cite{murakamiSc03} For PSC there is no net motion of charge; spin-up
electrons move in a given direction, while spin-down electrons travel in the
opposite one. This effect can be due to one-photon absorption of linearly
polarized light by a semiconductor, with filled valence bands and empty
conduction bands, illuminated by light with photon energy larger than the
energy gap. This phenomenon can be due to spin injection,\cite{malPRB03} Hall
Effects,\cite{sinovaPRB04} interference of two optical beams,\cite{bhatPRL00,
najmaiePRB03} or one photon absorption of linearly polarized
light\cite{bhatPRL05}. The last effect has been observed in gallium arsenide
(GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{fig1}
    \caption{(color online) Top (top panel) and side (bottom panel) views of
the Up structure along with the Cartesian $x$, $y$, and $z$ directions. The
dark (light) spheres are the C (H) atoms. The primitive hexagonal unit cell is
also shown.}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=5cm]{fig2}
    \caption{(color online) Top (top panel) and side (bottom panel) views of
the Alt structure along with the Cartesian $x$, $y$, and$ z$ directions. The
dark (light) spheres are the C (H) atoms. The primitive rectangular unit cell
is also shown. }
    \label{fig:alt-struc}
\end{figure}

@@

In this section, we  summarize the theoretical approach, involved in the
calculation of the spin velocity injection (SVI) resulting from the pure spin
current (PSC).

@@

in this work. 
To include the coherence, we follow Ref.~\onlinecite{nastosPRB07}

@@

single particle density matrix 
$\hat{\rho}({\mathbf k};t)$,  
leading to

@@

&\frac{\partial \rho_{cc'}({\mathbf k;t})}{\partial t} =

@@

states, and we take $\epsilon\to 0$ at the end of the calculation. 
Since the 
spin-splitting of the valence ($v$) bands is very small, we neglect it throughout
this work,\cite{nastosPRB07} and then
$\rho_{vv'}({\mathbf k};t)= 
\rho_{cv}({\mathbf k};t) =0$. 
 The matrix elements of any operator ${\cal O}$ are

@@

n{\mathbf k}| = 1$, where $n$ goes over all $v$ and $c$ states.
Therefore, using  Eqs. \eqref{z.4}

@@

–identical to Eq. (3) of Ref. \onlinecite{bhatPRL05} derived using the

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we calculate the magnitude
of the electron velocity along the plane of the structure, with the spin
polarized along $\hat{\mathbf{b}}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
\left(\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)\right)^{2}\ +
\left(\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)\right)^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\sigma^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
We also define two special angles
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular to the
incoming light polarization direction, respectively. 
The subscript $\sigma^\mathrm{b}$
denotes the spin along $\hat{\mathbf{b}}$.

@@

Fixing the calculated velocity along $\mathrm{a}=x$ or $\mathrm{a}=y$, we
define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
\left(\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)\right)^{2} +
\left(\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)\right)^{2} +
\left(\mathcal{V}^{\mathrm{az}}(\omega,\alpha)\right)^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented in the $xyz$ system of
coordinates along the polar angle, 
\begin{equation}
\theta_{\mathrm{a}}  (\omega,\alpha) = 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{equation}
and the azimuthal angle
\begin{equation}
\varphi_{\mathrm{a}} (\omega,\alpha) =
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{equation}

@@

We present the calculated results of
$\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ and
$\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ for the Up  and
Alt structures, both noncentrosymmetric 2D carbon systems with
50\% hydrogenation, which are differently structurally arranged. We remind that
the Up structure has hydrogen atoms only on the upper side of the carbon sheet,
while the Alt structure has alternating hydrogen atoms on the upper and bottom
sides. We take the carbon lattice to be along the $xy$ plane for both
structures, and the carbon-hydrogen bonds are
perpendicular to  $xz$ plane for the Up structure
(Fig. \ref{fig:up-struc}),
 and off the normal for the Alt structure 
(Fig. \ref{fig:alt-struc}). The coordinates for the Up and Alt unit cells of the
structures are presented in Tables \ref{tab:up-unitcell} and 
\ref{tab:alt-unitcell}, respectively.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{fig3}
\caption{(color online)  Spin velocity injection
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ {\it vs.} photon energy
  $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The largest
  velocity are at the low energy regions of the spectra for the Alt and Up
  structures, becoming different from zero at the energy gap of each structure.
  In the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to
  the 3D case of CdSe and GaAs, shown at the bottom panel.}
\label{fig:vab-str-comp}
\end{figure}

In Fig. \ref{fig:vab-str-comp}, we show $\mathcal{V}^{\mathrm{ab}}
(\omega,\alpha)$ {\it vs.} $\hbar\omega$ for the 
velocity and spin
directions
$\hat{\mathbf{a}}$ and $\hat{\mathbf{b}}$, and for the angle $\alpha$, at which
the signal is maximized, for the Up  and Alt structures, and for CdSe and GaAs
bulk systems, shown for comparison. As expected from
Eq.~\eqref{eq:mu}, 
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ starts rising 
from zero right at the
corresponding energy gap of each system. For the 2D structures considered, the
spectrum contains two narrow energy regions with strong response, while for the
bulk systems, the spectra covers a rather wide energy range, but with a much
weaker response. For the Up structure, at $\mathrm{a}\mathrm{b}=yz$ and
$\alpha=35^\circ$ the response is maximized, which means that an incoming light
with its electric field polarized at 35$^\circ$ from the $x$ direction will
induce electrons to move along $y$ (parallel to the surface), with their spin
polarized along $z$ (perpendicular to the surface). Right at the energy onset, ${\cal
V}^{yz}(\omega,\alpha)=668$\,Km/s, remains almost constant for 65 meV, and then
decreases to zero. A second region with high velocity is above 1.946 eV with
two, opposite in sign, maxima of the speed: ${\cal
V}^{yz}(\omega,\alpha)=266.3$ Km/s at $\hbar\omega=1.954$\,eV, and ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)=-241.4$ Km/s at
$\hbar\omega=1.958$\,eV; a positive (negative) ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ means that the electrons move
parallel (antiparallel) to the electric field. Likewise, for the Alt structure,
we also find that $\mathrm{a}\mathrm{b}=yz$ and $\alpha=150^\circ$ maximizes
the response, where two extreme values of ${\cal V}^{\mathrm{y}\mathrm{z}}$ are
found, one at  $\hbar\omega=0.720$\,eV of ${\cal
V}^{\mathrm{y}\mathrm{z}}=-711.9$ Km/s, and the other at
$\hbar\omega=0.911$\,eV of ${\cal V}^{\mathrm{y}\mathrm{z}}=-330.6$ Km/s.

For the bulk structures, we calculate ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$
from Eq.~\eqref{eq:vab-w} by simply using $\boldsymbol{\mu}_{\mathrm{max}}$.
For CdSe, we find that for $\hbar\omega=0.844$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to \mu^{zzzz}$, and ${\cal
V}^{\mathrm{z}\mathrm{z}}(\omega)=-59.0$ Km/s, and for GaAs at
$\hbar\omega=2.324$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to\mu^{\mathrm{aaaa}}$ and ${\cal
V}^{\mathrm{a}\mathrm{a}}(\omega,\alpha)=-28.7$\,Km/s, with $\mathrm{a}=x,y,z$.
For these bulk semiconductors, the $x$, $y$, and $z$ axis are taken along the
standard cubic unit cell directions, [100], [010], and [001], respectively. In
Table \ref{tab:vab-str-comp}, we compare ${\cal
V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ for the 2D structures considered and
bulk crystals. We stress that, as shown in the figure, the 2D structures
have maxima in ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$  higher than
for the bulk crystals by more than order of magnitude. 
In particular, the Alt structure demonstrates a 
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$ 
about 12 times larger than that
of CdSe and GaAs.

\begin{table}
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
System \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$}\\
\cline{5-6}
& type & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
Up  & 2D   & 35    & 0.084  & $\mathrm{yz}$ &  660.5 \\
      &      &       & 1.954  & $\mathrm{yz}$ &  266.3 \\
      &      &       & 1.958  & $\mathrm{yz}$ & -241.4 \\
Alt & 2D   & 150   & 0.720  & $\mathrm{yz}$ & -711.9 \\
      &      &       & 0.911  & $\mathrm{yz}$ & -330.6 \\
CdSe  & bulk & --    & 0.844  & $\mathrm{zz}$ &  -59.0 \\
GaAs  & bulk & --    & 2.324  & $\mathrm{xx}$ &  -28.7 \\
\hline
\end{tabular}
\caption{Comparison of the reported maximum values of
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ for the different structures and
their corresponding polarization angle $\alpha$ and energy $\hbar\omega$ . }
\label{tab:vab-str-comp}
\end{table}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{fig4}
\caption{(color online) For the Up structure, the top panel
shows ${\cal V}_{\sigma^\mathrm{z}}(\omega,\alpha)$ {\it vs.} $\hbar\omega$ and
$\alpha$, and the bottom panel shows
$\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ (right scale, red short-dashed line), and ${\cal
V}_{\sigma^\mathrm{z}}(\omega,\alpha)$ (left scale, black solid line), {\it
vs.} $\alpha$, for $\hbar\omega=0.084$\,eV, i.e. along the ridge shown in the
3D plot. }
\label{fig:up-vsz-w1}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{fig5}
\caption{(color online) For the Up structure, the top panel shows ${\cal
V}_{\sigma^\mathrm{z}} (\omega,\alpha)$ {\it vs.} $\hbar\omega$ and $\alpha$,
and the bottom panel shows $\gamma_{\sigma^\mathrm{z}}
(\omega,\alpha)$ 
(right scale, red short-dashed line), and ${\cal V}_{\sigma^\mathrm{z}}
(\omega,\alpha)$ (left scale, black solid line), {\it vs.} $\alpha$, for
$\hbar\omega=1.954$\,eV, i.e. along the highest ridge shown in the 3D plot. }
\label{fig:up-vsz-w2}
\end{figure}

@@

Here we calculated $\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ (Eq. 
\eqref{eq:vv-mag}) after fixing the electron velocity direction, $\hat{\mathbf{a}}$,
to the $x$ or $y$ direction along the surface of the Up and Alt structures,
and from Eqns. \eqref{eq:polar-ang} and \eqref{eq:azimuthal-ang}, 
we
determined the corresponding polar angle, $\theta_{\mathrm{a}}
(\omega,\alpha)$, and azimuthal angle, $\varphi_{\mathrm{a}}
(\omega,\alpha)$, of the resulting spin orientation.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{fig7}
\caption{(color online) For the Up structure we show the velocity
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$ (left scale, black solid line), the polar
angle $\theta_{\mathrm{a}} (\omega,\alpha)$ (right scale, red dashed  line), and  the
azimuthal angle $\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue dot-dashed 
line), {\it vs.} $\hbar\omega$, for $\alpha=35^\circ$, and $\mathrm{a}=x$ and
$\mathrm{a}=y$.}
\label{fig:up-vab-comp-rtp-1}
\end{figure}
